---
title: "Extracting + cleaning data for interim analysis of Dyna-CORE study"
authors: 
"Lara Puhlmann (lara.puhlmann@lir-mainz.de)" 
"Matthias Zerban (matthias.zerban@unimedizin-mainz.de)"  
"Jeroen Weermeijer (jeroen.weermeijer@kuleuven.be)"
"Haakon Engen"

date: "April 10th, 2020"
---

Formats data for interim analysis and includes plausibility checks

WARNING: Only run once! 

definition of Europe based on United Nations, Department of Economic and Social Affairs, Population Division (2019). World Population Prospects 2019, Online Edition. Rev. 1.


```{r setup, include=FALSE}

library(knitr)
require(foreign)
require(plyr)
require(dplyr)
require(stringr)
require(BBmisc)
require(formattable)
#require(Hmisc)
#require(corrplot)

# source functions to renames variables and occupation and occupational status in lists
source("rename.R") 
source("formatting.R")

# function to extract numeric part free response
numextract <- function(string){str_extract(string, "\\-*\\d+\\.*\\d*")} 

```

```{r Read and format data}
# load data - currently set up to work with owncloud. Data must have 171 columns!

#data_en <- read_csv("DynaCORE - the DynaMORE study on psychological responses to the Corona.csv", sep = ",", stringsAsFactors = FALSE)
#data_text <- read_csv("DynaCORE-C_text_answers.csv", sep = ",", stringsAsFactors = FALSE) # text version 

# user-specific path [change accordingly]:
data_en = read.csv("C:/Users/Matze/ownCloud/data/DynaCORE_C/DynaCORE - the DynaMORE study on psychological responses to the Corona.csv", sep = ",", stringsAsFactors = FALSE)
#data_en = read.csv("C:/Users/Nutzer/ownCloud/DynaCORE_C/DynaCORE - the DynaMORE study on psychological responses to the Corona.csv", sep = ",", stringsAsFactors = FALSE)
data_text = read.csv("C:/Users/Matze/ownCloud/data/DynaCORE_C/DynaCORE-C_text_answers.csv", sep = ",", stringsAsFactors = FALSE)
#data_text = read.csv("C:/Users/Nutzer/ownCloud/DynaCORE_C/DynaCORE-C_text_answers.csv", sep = ",", stringsAsFactors = FALSE)

data_en = rename(data_en) #rename variables
data_en$X.0 = data_en$occupation
data_en$X.00 = data_en$occupational.status
data_en = formatting(data_en) #group occupation + status in lists

data_en = data_en[which(!is.na(data_en$Respondent.ID)),] # remove row without respondent ID

# indicate missings in questionnaires
data_en[,c(68:154,156:167)] <- lapply(data_en[,c(68:154,156:167)], as.numeric) # questionnaires
data_en$missings <- rowSums(is.na(data_en[,c(68:154,156:167)]))

# indicate missing data in covariates:
data_en$missing.cov = NA

# 1) anyone away from their residence and not specifying the country or city of their current location
xx  = which(data_en$current.stay.out.of.town==1 & nchar(data_en$current.stay.out.of.town.country)==0 | data_en$current.stay.out.of.town==1 & nchar(data_en$current.stay.out.of.town.city)==0)
data_en$missing.cov[xx] = 1

# 2): missing symptom specification
xx = which(data_en$infection.test.status=="0" & is.na(as.numeric(data_en$symptom.severity)))
data_en$missing.cov[xx] = 1

```

```{r quality check income, include = FALSE}
i <- which.first(data_en$Respondent.ID[2]==data_text$Respondent.ID) -1
data_en$check_ID_text <- data_text$Respondent.ID[i:nrow(data_text)]
data_en$check_income_text <- data_text$What.is.your.approximate.average.annual.household.income..please.estimate.in.Euro..[i:nrow(data_text)]
test <- data_en[,c(ncol(data_en)-1, ncol(data_en))] #subset to check income variable
test$ID <- data_en$Respondent.ID
isTRUE(test$check_ID_text[nrow(test)]==test$ID[nrow(test)])
test$income <- data_en$household.income
table(test$income) 
#            1       10       11       12       13       14        2        3        4        5        6        7 
#3174      416      409      205      123       60      130      936      742      777     1273     2045     1392 
#  8        9 Response 
#750      422        1 

table(test$income[test$check_income_text=="â‚¬25,000-â‚¬49,999"]) 
#there are multiple levels coded to this response:
#    10    6 
#1  409 2045

```

```{r date & time}

#split month-day-year date + time col into two col, one with the date format, one with correct time
for(i in 1:length(data_en$Respondent.ID)){
  start = strsplit(data_en$Start.Date[i], " ")
  data_en$Start.Date[i] = start[[1]][1]
  data_en$Start.Time[i] = paste(start[[1]][2], start[[1]][3])
  
  end = strsplit(data_en$End.Date[i], " ")
  data_en$End.Date[i] = end[[1]][1]
  data_en$End.Time[i] = paste(end[[1]][2], end[[1]][3])
}

# dealing with different separators for date outputs (e.g. mm/dd/yyyy vs. mm.dd.yyyy).
data_en$Start.Date = gsub(".", "/", data_en$Start.Date, fixed=TRUE) #mm.dd.yyyy becomes mm/dd/yyyy

#convert month-day-year to year-month-day date, then to POSIXlt
data_en$Start.Date = as.Date(data_en$Start.Date, tryFormats = c("%m-%d-%Y", "%m/%d/%Y"), optional = FALSE)
data_en$End.Date = as.Date(data_en$End.Date, tryFormats = c("%m-%d-%Y", "%m/%d/%Y"), optional = FALSE)
data_en$Start.Date = as.POSIXlt(paste(data_en$Start.Date), tz = "Europe/Berlin", format="%Y-%m-%d")
data_en$End.Date = as.POSIXlt(paste(data_en$End.Date), tz = "Europe/Berlin", format="%Y-%m-%d")
data_en$Start.DateTime = as.POSIXlt(paste(data_en$Start.Date, data_en$Start.Time), tz = "Europe/Berlin", format="%Y-%m-%d %I:%M:%S %p")
data_en$End.DateTime = as.POSIXlt(paste(data_en$End.Date, data_en$End.Time), tz = "Europe/Berlin", format="%Y-%m-%d %I:%M:%S %p")

# #test
# data_en$Start.Date[4] #should give year/month/day GMT
# data_en$End.Date[4] #should give year/month/day GMT
# data_en$Start.DateTime[4] #should give year/month/day hour/minutes/seconds GMT
# data_en$End.DateTime[4] #should give year/month/day hour/minutes/seconds GMT

#completion time
data_en$completionTime = difftime(data_en$End.DateTime, data_en$Start.DateTime)
# test: data_en$completionTime[3] #gives time difference in mins
data_en$completionTime <- as.numeric(data_en$completionTime, units="secs")

```

```{r data up to 5000th complete response from European residents, include=FALSE}

Europe = c(2, 4, 11, 17, 18, 23, 28, 45, 48, 51, 60, 63, 64, 68, 70, 77, 80, 81, 86, 88, 98, 103, 104, 105, 111, 117, 119, 127, 132, 142, 143, 146, 147, 148, 154, 158, 162, 163, 168, 174, 175, 180, 191, 193)

data_en = data_en[order(data_en$End.DateTime),] # sort by completion date + time

data_en$complete.eu = NA #count of complete responses from European residents
data_en$complete.all = NA #count of complete responses overall
c = 0
c2 = 0
for(i in 1:length(data_en$complete.eu)){
  if (data_en$missings[i] == 0 && is.na(data_en$missing.cov[i]) && data_en$country.of.residence[i] %in% Europe){
  c = c+1
  data_en$complete.eu[i] = c
  } 
  if(data_en$missings[i] == 0 && is.na(data_en$missing.cov[i])) {
  c2 = c2+1
  data_en$complete.all[i] = c2
  }
}
data_en = data_en[1:which(data_en$complete.eu==5000),] # keep data up until the 5000th complete response
dim(data_en[data_en$missings == 0 &is.na(data_en$missing.cov)&data_en$country.of.residence %in% Europe,]) # should be 5000
dim(data_en[data_en$missings > 0|!is.na(data_en$missing.cov)|!(data_en$country.of.residence %in% Europe),]) # should be the remaining

data_eu = data_en[which(data_en$country.of.residence %in% Europe),] # exclude all non-European residents
data_ex = data_en[which(!(data_en$country.of.residence %in% Europe)),] # exclude all non-European residents

#incomplete responses:
dim(data_eu[data_eu$missings > 0|!is.na(data_eu$missing.cov),])
dim(data_ex[data_ex$missings > 0|!is.na(data_ex$missing.cov),])
dim(data_eu[data_eu$missings == 0&is.na(data_eu$missing.cov),])
dim(data_ex[data_ex$missings == 0&is.na(data_ex$missing.cov),])
dim(data_ex[data_ex$missings > 0,]) # very high n of incomplete resp
write.csv(data_ex[,c(1:21, 52:179)], "data_outside_eu.csv") # truly, loads are incomplete

max(data_en$End.DateTime)
data_en$complete.eu[which(data_en$End.DateTime == max(data_en$End.DateTime))]
data_en$End.DateTime[length(data_en$Respondent.ID)] # pull date

data_all = data_en # all data up to 'pull date'
# data_en = data_en[which(data_en$country.of.residence %in% Europe),] # exclude all non-European residents

```

```{r clean "out of town" response in char format, include = FALSE}

# set "out of town" response of people naming a different country than their country of residence to yes
xx = which(data_en$current.stay.out.of.town.country[which(data_en$current.stay.out.of.town=="2" & data_en$current.stay.out.of.town.country !="")] != data_en$country.of.residence[which(data_en$current.stay.out.of.town=="2" & data_en$current.stay.out.of.town.country !="")])

data_en$current.stay.out.of.town[which(data_en$current.stay.out.of.town=="2" & data_en$current.stay.out.of.town.country !="")][xx] = "1"

# set responses of people not out of town and naming the same as their country of residence to empty
length(which(data_en$current.stay.out.of.town.country[which(data_en$current.stay.out.of.town=="2")] != ""))
length(which(data_en$current.stay.out.of.town.country[which(data_en$current.stay.out.of.town=="2"&!is.na(data_en$complete.eu))] != ""))

data_en$current.stay.out.of.town.country[which(data_en$current.stay.out.of.town=="2" & data_en$current.stay.out.of.town.country != "")] <- ""

data_en$current.location <- ifelse(data_en$current.stay.out.of.town == '1', data_en$current.stay.out.of.town.country, data_en$country.of.residence)
 
# test
data_en$country.of.residence[which(data_en$current.stay.out.of.town==1& data_en$missings ==0)][17] 
data_en$current.stay.out.of.town.country[which(data_en$current.stay.out.of.town==1& data_en$missings ==0)][17] 
data_en$current.location[which(data_en$current.stay.out.of.town==1& data_en$missings ==0)][17]

# how many are not currently in Europe?
length(which(!(data_en$current.stay.out.of.town.country %in% Europe) & data_en$current.stay.out.of.town.country !=""))
data_en$current.stay.out.of.town.country[which(!(data_en$current.stay.out.of.town.country %in% Europe) & data_en$current.stay.out.of.town.country !="")]
```

```{r  format covariates}
#data_en[,c(68:154,156:167)] <- lapply(data_en[,c(68:154,156:167)], as.numeric) # questionnaires
data_en[,c(1:2, 10:12,14:16, 18:19, 53:54, 58, 60:61,64)] <- lapply(data_en[,c(1:2, 10:12,14:16, 18:19, 53:54, 58, 60:61,64)], as.factor)# covariates

data_en$health.status = as.numeric(data_en$health.status)
data_en$household.income = factor(data_en$household.income, order = TRUE)
data_en$symptom.severity = factor(data_en$symptom.severity, order = TRUE)

data_en$people.in.household.under.18 = as.numeric(data_en$people.in.household.under.18)
data_en$opinion.about.authorities.measures = as.numeric(data_en$opinion.about.authorities.measures)
data_en$adherence.to.recommended.procedures = as.numeric(data_en$adherence.to.recommended.procedures)

# people.in.household as continuous
data_en$people.in.household.cont = as.numeric(data_en$people.in.household)-1 # factor 0, meaning more than 6, was recoded to 1
xx = as.numeric(numextract(data_en$X.28))
xx[which(xx==0)] = 1
data_en$people.in.household.cont[which(data_en$people.in.household.cont == 0)] = xx[!is.na(xx)]
data_en$people.in.household.cont[which(data_en$people.in.household.cont == 3)] = 3.5
data_en$people.in.household.cont[which(data_en$people.in.household.cont == 4)] = 5.5


# date of Corona test
data_en$infection.test.status.date = gsub(".", "/", data_en$infection.test.status.date, fixed=TRUE)
data_en$infection.test.status.date[which(nchar(data_en$infection.test.status.date)<5)]=NA # set all that are not a date to NA
data_en$infection.test.status.date = as.Date(data_en$infection.test.status.date, tryFormats = c("%m-%d-%Y", "%m/%d/%Y"), optional = FALSE) # convert month-day-year to year-month-day date
data_en$infection.test.status.date = as.POSIXlt(paste(data_en$infection.test.status.date), tz = "Europe/Berlin", format="%Y-%m-%d") # date as POSIXlt
```

```{r covariates: plausibility checks & cleaning, include = FALSE}
# remaining incomplete covariates among those without missings - should all be 0
data_en.complete = data_en[which(!is.na(data_en$complete.eu)),]
#length(which(data_en.complete$current.stay.out.of.town==1 & data_en.complete$current.stay.out.of.town.city==""))
length(which(data_en.complete$people.in.household==0 & nchar(data_en.complete$X.28)==0))
length(which(data_en.complete$infection.test.status==0 & is.na(as.numeric(data_en.complete$symptom.severity))))


###### age #####

data_en$age.fulltext = data_en$age
data_en$age = gsub("o", "0", data_en$age) #account for people accidently typing 'o' instead of '0'

# extract the numeric component from age response
data_en$age = numextract(data_en$age)
data_en$age = as.numeric(data_en$age)
data_en$age[which(data_en$age > 100)] # 111 seems possible, so keep
length(which(is.na(data_en$age))) # none

# if any ages are 0, this could be due to a leading o that was substituted above. Check fulltext:
data_en$age.fulltext[which(data_en$age==0)] # -> one person, who says to be 41 soon
data_en$age[which(data_en$age==0)] = 41 # recode this person:

# check invalid age responses 
which(data_en$age < 18) # four
which(data_en$age[which(!is.na(data_en$complete.eu))] < 18)

which(is.na(data_en$age[which(!is.na(data_en$complete.eu))])) # none

# exclude those resposes
data_en$Respondent.ID[which(data_en$age < 18)]<- NA

data_en$Respondent.ID[which(is.na(data_en$age))]<- NA
xx = which(is.na(data_en$Respondent.ID))
if(length(xx)>0){data_en = data_en[-xx,]}

###### education #####

data_en$years.of.education.fulltext = data_en$years.of.education
data_en$years.of.education = numextract(data_en$years.of.education) # extract the numeric component of response
data_en$years.of.education = as.numeric(data_en$years.of.education)
# if people fail to add their years of education to a total, they may end up with too low of a value. "years.of.education.fulltext" includes the full answer for years.of.education for anyone with less than 10 years

length(which(data_en$years.of.education > data_en$age)) # indicate people with more years of education than age
length(which(data_en$years.of.education[which(!is.na(data_en$complete.eu))] > data_en$age[which(!is.na(data_en$complete.eu))]))

for(i in 1:length(data_en$Respondent.ID)){
  if (!is.na(data_en$years.of.education[i])) {
    if(data_en$years.of.education[i] > data_en$age[i]){
      data_en$years.of.education[i] = NA
    }
    if(!is.na(data_en$years.of.education[i]) && data_en$years.of.education[i] > 10){
      data_en$years.of.education.fulltext[i] = NA
    }
  }
}

# check fulltest answers to make sure this was not due to typos or nor summing the total years of 
data_en$years.of.education.fulltext[which(!is.na(data_en$years.of.education.fulltext))] # all plausible

####  inconsistent responses ####

# count cases where more/same as total nr of people in household are underage
xx = which(data_en$people.in.household.cont+0.5<=data_en$people.in.household.under.18)
xx = which(data_en$people.in.household.cont[which(!is.na(data_en$complete.eu))]+0.5<=data_en$people.in.household.under.18[which(!is.na(data_en$complete.eu))])

# set symptom ratings of those not being COVID positive to NA
xx = which(data_en$symptom.severity[which(data_en$infection.test.status==1)] > 1) # symptom severity above 1 although saying they were not tested pos
#xx = which(data_en.complete$symptom.severity[which(data_en.complete$infection.test.status==1)] > 1)
length(which(data_en$symptom.severity[which(data_en$infection.test.status==1)] != ""))
data_en$symptom.severity[which(data_en$infection.test.status==1)] <- ""


### occupation

# variable with occupation of people with only one occupation
data_en$occupation.if.only.one <- as.character(data_en$occupation)
data_en$occupation.if.only.one <- as.numeric(data_en$occupation.if.only.one)
# variable with occupational status of people with only one indicated
data_en$occupational.status.if.only.one <- as.character(data_en$occupational.status)
data_en$occupational.status.if.only.one <- as.numeric(data_en$occupational.status.if.only.one)

#data_en$occupational.status[sapply(data_en$occupational.status, function(x)  "1" %in% x)]

### set cases with mismatch in occupation to NA
data_en$not.working.12 <- lapply(data_en$occupation, function(ch) grep("16", ch))
data_en$not.working.12[sapply(data_en$not.working.12, function(x) length(x)==0)] <- NA

employed = c("1", "2", "3", "4") # Note that this also grabs 10, 11, 12
not.employed = c("7", "8", "9", "10", "11", "12") #parental leave, studying, sick leave, unemployment w/ or w/o benefits, retired
data_en$employed.13 <- lapply(data_en$occupational.status, function(ch) grep(paste(employed, collapse="|"), ch))
data_en$not.employed.13 <- lapply(data_en$occupational.status, function(ch) grep(paste(not.employed, collapse="|"), ch))
data_en$employed.13[sapply(data_en$employed.13, function(x) length(x)==0)] <- NA
data_en$not.employed.13[sapply(data_en$not.employed.13, function(x) length(x)==0)] <- NA
data_en$employed.13[which(!is.na(data_en$not.employed.13))] <- NA # also set to NA if there is 

# find people that are unemployed in 12 but working in 13
index = which(!is.na(data_en$not.working.12)) %in% which(!is.na(data_en$employed.13))
xx = which(!is.na(data_en$not.working.12))[index]
data_en$occupation[xx] <- NA
data_en$occupational.status[xx] <- NA

# # add 'not working' to occupation of all individuals listing forms of not working in 13
# index = which(is.na(data_en$not.working.12)) %in% which(!is.na(data_en$not.employed.13))
# xx = which(is.na(data_en$not.working.12))[index]
# data_en$occupation[xx][[1]][length(data_en$occupation[xx][[1]])+1] <- 16

# drop empty levels
data_en$gender = droplevels(data_en$gender)
data_en$nationality = droplevels(data_en$nationality)
data_en$country.of.residence = droplevels(data_en$country.of.residence)
data_en$current.stay.out.of.town = droplevels(data_en$current.stay.out.of.town)
data_en$current.stay.out.of.town.country = droplevels(data_en$current.stay.out.of.town.country)
data_en$infection.test.status = droplevels(data_en$infection.test.status)
data_en$survey.language = droplevels(data_en$survey.language)
data_en$relationship.status = droplevels(data_en$relationship.status)
data_en$people.in.household = droplevels(data_en$people.in.household)
data_en$risk.group = droplevels(data_en$risk.group)
data_en$quarantine.status = droplevels(data_en$quarantine.status)
data_en$household.income = droplevels(data_en$household.income)
```

```{r data checks, include=FALSE}

# indicate individuals who report COVID symptoms but in stressor exposure said this situation did not happen
data_en$symptom.inconsistency = NA
data_en$symptom.inconsistency[which(data_en$symptom.severity >0 & data_en$CE_01 == 0)] = 1
sum(data_en$symptom.inconsistency, na.rm = T) # -> none

# indicate individuals who report being in a risk group but in stressor exposure said to risk group "this situation did not happen"
data_en$risk.group.inconsistency = NA
data_en$risk.group.inconsistency[which(data_en$risk.group == 1 & data_en$CE_04 == 0)] = 1
sum(data_en$risk.group.inconsistency, na.rm = T)

## find people who provide a stressor rating > 0 in the free answers but fail to describe the stressor
length(which(data_en$CE_30>0 & nchar(data_en$CE_30_text) == 0))
length(which(data_en.complete$CE_30>0 & nchar(data_en.complete$CE_30_text) == 0))

length(which(data_en$GE_12>0 & nchar(data_en$GE_12_text) == 0))
length(which(data_en.complete$GE_12>0 & nchar(data_en.complete$GE_12_text) == 0))

# number of responses per country
sort( table(unlist(data_en$country.of.residence)),decreasing=TRUE)

# frequency table of 10 most frequent countries
sort( table(unlist(data_en$country.of.residence)),decreasing=TRUE)[1:10]

```

```{r old income coding}
 
data_en$household.income.old = factor(data_en$household.income, order = TRUE)
test <- data_en[c("Respondent.ID","check_ID_text","household.income","household.income.old", "check_income_text")] #to check corresponding factors
data_en$household.income.old <- mapvalues(data_en$household.income.old,c(1,2,3,4,5,6,7,8,9,10,11,12), c(1,1,1,1,1,6,7,8,9,6,11,12))
```

```{r restructure questionnaire variables}

#data_en[,c(68:154,156:167)] <- lapply(data_en[,c(68:154,156:167)], as.numeric)
#data_en <- data_en[data_en$missings == 0,] # remove any cases with missings

# Mental Health Problems [P]
data_en$CM_07 <- 5 - data_en$CM_07
term <- "CM"
GHQ <- grep(term, names(data_en))
GHQrecode <- function(x){recode(x, '1'=0L, '2'=1L, '3'=2L, '4'=3L)}
data_en[GHQ] <- lapply(data_en[GHQ], GHQrecode)
GHQ <- GHQ[1:12]
data_en$P <- rowSums(data_en[GHQ])

# percieved social support [PSS]
term <- "H2_"
PSSindex <- grep(term, names(data_en))
PSSindex <- PSSindex[1:7]
data_en$PSS <- rowSums(data_en[PSSindex])

# Optimism:
data_en$OPT <- as.numeric(data_en$H3_01)

# Perceived general self efficacy
term <-"H4_"
GSEindex <- grep(term, names(data_en))
data_en$GSE <- rowSums(data_en[GSEindex])

# self-percieved resilience (BRS)
term <- "H5_"
BRS <- grep(term, names(data_en))
BRSrec <- c("H5_02", "H5_04", "H5_06")
data_en[,BRSrec] <- 6 - data_en[,BRSrec]
data_en$REC <- rowMeans(data_en[BRS])

# BFI Neuroticism
data_en$H6_01 <- 6 - data_en$H6_01
term <- "H6_"
BFI <- grep(term, names(data_en))
BFIrecode <- function(x){recode(x, '1'=-2L, '2'=-1L, '3'=0L, '4'=1L,'5'=2L)}
data_en[BFI] <- lapply(data_en[BFI], BFIrecode)
data_en$NEU <- rowSums(data_en[BFI])

# Behavioral Coping style
term <- "COPE"
COPE <- grep(term, names(data_en))
COPE <- COPE[c(1:5,7:9)]
data_en$BCS <- rowSums(data_en[COPE])

# CERQ
term <- "CERQ"
CERQ <- grep(term, names(data_en))
data_en$CERQSum <- rowSums(data_en[CERQ])

# Positive Appraisal Style:
PAS <- data_en[,c(CERQ, 106, 110 )]
PAS[,c("H1_COPE_18","H1_COPE_28")] <- PAS[,c("H1_COPE_18","H1_COPE_28")]*5/4 #rescale
data_en$PAS <- rowMeans(PAS)

# Corona specific appraisal
term <- "H1_Cor_"
PAC <- grep(term, names(data_en))
data_en$PAC <- rowSums(data_en[PAC])

# summarise GHQ
length(data_en$P[which(!is.na(data_en$complete.all))])
mean(data_en$P[which(!is.na(data_en$complete.all))])
sd(data_en$P[which(!is.na(data_en$complete.all))])

length(data_en$P[which(is.na(data_en$complete.all))])
length(data_en$P[which(is.na(data_en$complete.all)&!is.na(data_en$P))])
mean(data_en$P[which(is.na(data_en$complete.all))], na.rm = T)
sd(data_en$P[which(is.na(data_en$complete.all))], na.rm = T)

data_en = data_en[which(data_en$country.of.residence %in% Europe),] # exclude all non-European residents

length(data_en$P[which(is.na(data_en$complete.eu))])
length(data_en$P[which(is.na(data_en$complete.eu)&!is.na(data_en$P))])
mean(data_en$P[which(is.na(data_en$complete.all))], na.rm = T)
sd(data_en$P[which(is.na(data_en$complete.all))], na.rm = T)

length(data_en$P[which(!is.na(data_en$complete.eu))])
mean(data_en$P[which(!is.na(data_en$complete.eu))])
sd(data_en$P[which(!is.na(data_en$complete.eu))])

```

```{r calculation of stressors}

# SCM = stressor count method
# SSM = stressor severity method

#Corona pandmeic related stressors:
term <- "CE_"
CE <- grep(term, names(data_en))
CE <- CE[1:30]
data_en$Es.SCM <- rowSums(data_en[CE] >0) #stressor count
data_en$Es.SSM <- rowSums(data_en[CE])/5 #weighted

#general stressors:
term <- "GE_"
GE <- grep(term, names(data_en))
GE <- GE[1:12]
data_en$Eg.SCM <- rowSums(data_en[GE] >0) #stressor count
data_en$Eg.SSM <- rowSums(data_en[GE])/5 #weighted

#combined:
Eall <- c(grep("GE_", names(data_en))[1:12], grep("CE_", names(data_en))[1:30])
data_en$Ec.SCM <- rowSums(data_en[Eall] >0) #stressor count
data_en$Ec.SSM <- rowSums(data_en[Eall])/5 #weighted

# test
which(data_en$Ec.SCM!=rowSums(data_en[ , c("Eg.SCM" ,"Es.SCM")])) # should be 0

```

```{r calculate SR Scores}

# stressor count method
m1 <- summary(lm(scale(P)~scale(Eg.SCM),data= data_en[!is.na(data_en$Eg.SCM),]))
data_en$SR_Eg.SCM[!is.na(data_en$Eg.SCM)] <-as.numeric(scale(resid(m1)))

m2 <- summary(lm(scale(P)~scale(Es.SCM),data= data_en[!is.na(data_en$Es.SCM),]))
data_en$SR_Es.SCM[!is.na(data_en$Es.SCM)] <-as.numeric(scale(resid(m2)))

m3 <- summary(lm(scale(P)~scale(Ec.SCM),data= data_en[!is.na(data_en$Ec.SCM),]))
data_en$SR_c.SCM[!is.na(data_en$Ec.SCM)] <-as.numeric(scale(resid(m3)))

# stressor severity method
m4 <- summary(lm(scale(P)~scale(Eg.SSM),data= data_en[!is.na(data_en$Eg.SSM),]))
data_en$SR_Eg.SSM[!is.na(data_en$Eg.SSM)] <-as.numeric(scale(resid(m4)))

m5 <- summary(lm(scale(P)~scale(Es.SSM),data= data_en[!is.na(data_en$Es.SSM),]))
data_en$SR_Es.SSM[!is.na(data_en$Es.SSM)] <-as.numeric(scale(resid(m5)))

m6 <- summary(lm(scale(P)~scale(Ec.SSM),data= data_en[!is.na(data_en$Ec.SSM),]))
data_en$SR_c.SSM[!is.na(data_en$Ec.SSM)] <-as.numeric(scale(resid(m6)))

```

```{r generate subgroupe, include = False}

# subjects in Europe
xx = which(data_en$current.location %in% Europe)
data_en$in.eu = 0
data_en$in.eu[xx] = 1
data_en$in.eu = as.factor(data_en$in.eu)

# index people with potentially precarious job conditions: freelancer, self-employed, temp contract, unemployed, by excluding everyone with a stable status
stable.occupational.status = c("1", "3", "7", "8", "11", "12")
insecure.occulational.status = c("2", "4", "5", "6", "9", "10")

index.stable.occupational.status <- lapply(data_en$occupational.status, function(ch) grep(paste(stable.occupational.status, collapse="|"), ch))
index.stable.occupational.status[sapply(index.stable.occupational.status, function(x) length(x)==0)] <- NA

index.insecure.occupational.status <- lapply(data_en$occupational.status, function(ch) grep(paste(insecure.occulational.status, collapse="|"), ch))
index.insecure.occupational.status[sapply(index.insecure.occupational.status, function(x) length(x)==0)] <- NA

data_en$unstable.occupational.status = NA 
data_en$unstable.occupational.status[which(!is.na(index.stable.occupational.status))] = 0 # 0 = no, stable status
data_en$unstable.occupational.status[which(is.na(index.stable.occupational.status)&!is.na(index.insecure.occupational.status))] = 1 # 1 = yes, unstable status (and no additional stable status)

# index people in Germany
data_en$Germany = "no"
data_en$Germany[which(data_en$country.of.residence=="68")] = "yes"
length(data_en$Germany[which(data_en$Germany== 1& !is.na(data_en$complete.eu))])/length(data_en$Germany[which(!is.na(data_en$complete.eu))])
data_en$Germany = as.factor(data_en$Germany)

# index people with permanent contract
data_en$permanent_contract = "no"
contract = c("1", "3")
index.contract <- sapply(data_en$occupational.status, function(ch) contract %in% ch)
data_en$permanent_contract[which(index.contract[1,] > 0 |index.contract[2,] > 0)]="yes"
# #test
# data_en$occupational.status[which(data_en$permanent_contract==1)][1000:1100]
# data_en$occupational.status[which(data_en$permanent_contract==0)][1000:1100]
length(data_en$occupational.status[which(data_en$permanent_contract== "no" & !is.na(data_en$complete.eu))])
length(data_en$occupational.status[which(data_en$permanent_contract== "yes" & !is.na(data_en$complete.eu))])
length(data_en$occupational.status[which(data_en$permanent_contract== "yes" & !is.na(data_en$complete.eu))])/length(data_en$occupational.status[which(!is.na(data_en$complete.eu))])
data_en$permanent_contract = as.factor(data_en$permanent_contract)

# index people in academia
data_en$academia = "no"
academia = c("1", "2")
index.academia <- sapply(data_en$occupation, function(ch) academia %in% ch)
#data_en$occupational.status[which(index.academia[1,] > 0 |index.academia[2,] > 0)]
data_en$academia[which(index.academia[1,] > 0 |index.academia[2,] > 0)]="yes"
# #test
# data_en$occupation[which(data_en$academia=="yes")][1000:1100]
# data_en$occupation[which(data_en$academia=="no")][1000:1100]
length(data_en$occupation[which(data_en$academia== "no"& !is.na(data_en$complete.eu))])
length(data_en$occupation[which(data_en$academia== "yes" & !is.na(data_en$complete.eu))])/length(data_en$occupation[which(!is.na(data_en$complete.eu))])
data_en$academia = as.factor(data_en$academia)


# GHQ in academia
mean(data_en$P[which(data_en$academia== "no"& !is.na(data_en$complete.eu))])
sd(data_en$P[which(data_en$academia== "no"& !is.na(data_en$complete.eu))])
mean(data_en$P[which(data_en$academia== "yes"& !is.na(data_en$complete.eu))])
sd(data_en$P[which(data_en$academia== "yes" & !is.na(data_en$complete.eu))])

# psychiatric diagnoses
#ID.diagnosed <- read.csv("C:/Users/Matze/ownCloud/data/DynaCORE_C/included_psych_diagnosis.txt")
ID.diagnosed <- read.csv("C:/Users/Nutzer/ownCloud/DynaCORE_C/included_psych_diagnosis.txt")
names(ID.diagnosed)[1] <- "ID"
data_en$psychiatric.diagnose <- NA
#data_en$psychiatric.diagnose[which(ID.diagnosed$ID %in% data_en$Respondent.ID)] <- 1
data_en$psychiatric.diagnose[which(data_en$Respondent.ID %in% ID.diagnosed$ID)] <- 1

# newly included people with mental health conditions
mental.health = data.frame(ID = data_en$Respondent.ID[which(nchar(data_en$mental.health.details)>0)],
                           diagnosed.condition =  data_en$mental.health.details[which(nchar(data_en$mental.health.details)>0)],
                           complete = data_en$complete.eu[which(nchar(data_en$mental.health.details)>0)]
)
mental.health$complete[which(!is.na(mental.health$complete))] = 1
which(!(data_en$Respondent.ID[data_en$diagnosed.mental.health==0] %in% mental.health$ID)) # all included
length(which(mental.health$complete==1))

length(which(nchar(data_en$mental.health.details)>0))
length(which(data_en$diagnosed.mental.health==0)) # all who said yes specified their condition

#mental.health.old = read.csv("C:/Users/Matze/ownCloud/data/DynaCORE_C/mental.health.csv")
mental.health.old = read.csv("C:/Users/Nutzer/ownCloud/DynaCORE_C/mental.health.csv")

# new variables
mental.health$new = 0
mental.health$new[which(!(mental.health$ID %in% mental.health.old$ID))] = 1
sum(mental.health$new)

length(mental.health.old$ID[which(!(mental.health.old$ID %in% mental.health$ID))])

# not included anymore
mental.health.old$old = 0
mental.health.old$old[which(!(mental.health.old$ID %in% mental.health$ID))] = 1
sum(mental.health.old$old)

# do the numbers add up?
length(which(mental.health.old$complete==1)) - sum(mental.health.old$old[which(mental.health.old$complete==1)]) + sum(mental.health$new[which(mental.health$complete==1)]) # same as current n of mental health cases

length(which(mental.health$complete==1)) - sum(mental.health.old$old[which(mental.health.old$complete==1)]) + sum(mental.health$new[which(mental.health$complete==1)])

length(mental.health.old$ID[which(mental.health.old$old==1)])

mental.health.excluded.IDs = mental.health.old[which(mental.health.old$old==1),]

write.csv(mental.health, "mental.health.new.csv")
write.csv(mental.health.old, "mental.health.old.csv")
write.csv(mental.health.excluded.IDs, "mental.health.excluded.IDs.csv")

#answered yes to whether diagnosed but didn't specify:
# diagnosed.no.details = data.frame(ID = data_en$Respondent.ID[which(data_en$diagnosed.mental.health==0 & nchar(data_en$mental.health.details)==0)],
#                            diagnosed.condition = data_en$mental.health.details[which(data_en$diagnosed.mental.health==0 & nchar(data_en$mental.health.details)==0)],
#                            complete = data_en$complete.eu[which(data_en$diagnosed.mental.health==0 & nchar(data_en$mental.health.details)==0)]
# )
# diagnosed.no.details$complete[which(nchar(diagnosed.no.details$complete)>0)] <- 1
# 
# write.csv(diagnosed.no.details, "diagnosed.but.no.details.csv")

str(data_en$diagnosed.mental.health[which(!is.na(data_en$complete.eu))])
length(which(data_en$diagnosed.mental.health[which(!is.na(data_en$complete.eu))]==0))
length(which(data_en$diagnosed.mental.health[which(!is.na(data_en$complete.eu))]==""))

mean(data_en$P[which(!is.na(data_en$psychiatric.diagnose)& !is.na(data_en$complete.eu))])
sd(data_en$P[which(!is.na(data_en$psychiatric.diagnose)& !is.na(data_en$complete.eu))])
mean(data_en$P[which(is.na(data_en$psychiatric.diagnose)& !is.na(data_en$complete.eu))])
sd(data_en$P[which(is.na(data_en$psychiatric.diagnose)& !is.na(data_en$complete.eu))])
#test
which(!(ID.diagnosed$ID %in% data_en$Respondent.ID[which(data_en$diagnosed.mental.health==0)]))
ID.diagnosed$ID[which(!(ID.diagnosed$ID %in% data_en$Respondent.ID[which(data_en$diagnosed.mental.health==0)]))]
ID.diagnosed$ID[which(!(ID.diagnosed$ID %in% data_en$Respondent.ID))]

which((ID.diagnosed$ID %in% data_en$Respondent.ID[which(data_en$diagnosed.mental.health==0)]))

mean(data_en$P[which(data_en$diagnosed.mental.health==0& !is.na(data_en$complete.eu))])
sd(data_en$P[which(is.na(data_en$psychiatric.diagnose)& !is.na(data_en$complete.eu))])

mean(data_en$P[which(data_en$diagnosed.mental.health==1& !is.na(data_en$complete.eu))])


```

```{r response variance and completion time, include = False}

# exclude subjects with no response variance (check block-wise for all questionnaires with more than 2 items)
var = matrix(NA, nrow = length(data_en$Respondent.ID), ncol = 8)
for (i in 1:nrow(data_en)){ 
  var[i,1] = (var(as.vector(as.matrix(data_en[i, GHQ])))) 
  var[i,2] = (var(as.vector(as.matrix(data_en[i, PSSindex])))) 
  var[i,3] = (var(as.vector(as.matrix(data_en[i, GSEindex])))) #object asku not found
  var[i,4] = (var(as.vector(as.matrix(data_en[i, BRS])))) 
  var[i,5] = (var(as.vector(as.matrix(data_en[i, COPE])))) 
  var[i,6] = (var(as.vector(as.matrix(data_en[i, CERQ])))) 
  var[i,7] = (var(as.vector(as.matrix(data_en[i, CE])))) 
  var[i,8] = (var(as.vector(as.matrix(data_en[i, GE])))) 
}

data_en$response_variance = rowSums(var)
# exclude subject with 0 response variance
data_en$Respondent.ID[which(data_en$response_variance == 0)] # none

# distributions
hist(data_en$response_variance) # distribution of response variance - looks good
hist(data_en$age) # distribution of age
hist(data_en$completionTime/60) # distribution of completion time - some outliers
hist(data_en$completionTime[which(data_en$completionTime/60<1440)]/60) # distribution of completion time under 1 day
hist(data_en$completionTime[which(data_en$completionTime/60<60)]/60) # distribution of completion time under 1 hour
length(which(data_en$completionTime/60>1440)) # longer than 1 day
length(which(data_en$completionTime/60>2880)) # longer than 2 days

# very short completion time
which(is.na(data_en$completionTime[which(!is.na(data_en$complete.eu))]))
min(data_en$completionTime[which(!is.na(data_en$complete.eu))], na.rm = T)
# in complete sample at least 5.75 min, which is plausible


```

```{r first tables - perhaps move to analysis script?}
require(stargazer)
require(tableone)
require(kableExtra)
require(pander)

# full sample

# correlation of stressors
correlation.stressors <- cor(data_en[,c("Eg.SCM","Es.SCM","Ec.SCM", "Eg.SSM", "Es.SSM", "Ec.SSM")], use = "pairwise.complete.obs", method = "pearson")
correlation.P <- cor(data_en[,c("P", "Eg.SCM","Es.SCM","Ec.SCM", "Eg.SSM", "Es.SSM", "Ec.SSM")], use = "pairwise.complete.obs", method = "pearson")

stargazer(correlation.P[,1], title="Correlations of 'P' with stressors", type = "text", out="cor_P_stressors.txt")
#writeLines(capture.output(stargazer(correlation.P[,1], title="Correlations of 'P' with stressors", type = "text", out="correlation.P.txt")), "cor_P_stressors.txt")

stargazer(correlation.stressors, title="Correlation among stressors (Pearson r)", type = "text", out="correlation.stressors.txt")
#writeLines(capture.output(stargazer(correlation.stressors, title="Correlation among stressors", type = "html", out="correlation.P.html")), "cor_stressors.html")

# count stressor frequency in complete sample
data_en.complete = data_en[which(!is.na(data_en$complete.eu)),]
stressors.df = data.frame(names = c("a", "b", "c"),
                          stressors = c(colSums(data_en.complete[CE] >0), colSums(data_en.complete[GE] >0)))

# subsample
stressors.df = data.frame(Category = c("a", "b", "c"),
                          Count = c(colSums(data_en.complete[CE] >0), colSums(data_en.complete[GE] >0)),
                          Frequency = c(round(colSums(data_en.complete[CE] >0)/length(data_en.complete$CE_01)*100,2), round(colSums(data_en.complete[GE] >0)/length(data_en.complete$CE_01)*100,2)),
                          severity = c(round(apply(data_en.complete[CE],2,mean),2), round(apply(data_en.complete[GE],2,mean),2))
                          )
write.csv(stressors.df, "stressor_prevalence.csv")


# table one / table 1

demovar <- c("age", "gender", "nationality","country.of.residence", "current.stay.out.of.town",   "years.of.education", "occupational.status",
             "household.income.old", "relationship.status", "people.in.household", "diagnosed.mental.health",
             "risk.group", "infection.test.status", "quarantine.status", "opinion.about.authorities.measures",
             "adherence.to.recommended.procedures", "health.status", "people.in.household.under.18", "occupation", "occupational.status", "X.0", "X", "X.1", "X.2", "X.3", "X.4", "X.5", "X.6", "X.7", "X.8", "X.9" , "X.10" , "X.11" , "X.12" , "X.13" , "X.14" , "X.15" , "X.16" , "X.00" , "X.17" , "X.18" , "X.19" , "X.20" , "X.21" , "X.22" , "X.23" , "X.24" , "X.25" , "X.26" ,"X.27", "survey.language")
demog2 <- data_en[which(!is.na(data_en$complete.eu)), demovar]
demog <- data_en[which(is.na(data_en$complete.eu)), demovar]

# only incude common countries of residence

sort( table(unlist(demog$country.of.residence)),decreasing=TRUE)
# countries of res with N > 10:  68   88   18  142  127   80  175  168  193   64   60   11  174   51  132  143   63

#name most frequent countries
demog$country.of.residence = revalue(demog$country.of.residence,
      c("68" = "Germany", "142" = "Poland", "18" = "Belgium", "88" = "Italy", "127" = "Netherlands", "80" = "Hungary", "168" = "Spain", "175" = "Switzerland", "193" = "United Kingdom", "64" = "France", "60" = "Estonia", "11" = "Austria", "174" = "Sweden", "51" = "Denmark", "132" = "Norway", "63" = "Finland"))

frequent.countries = c("Germany", "Poland", "Belgium", "Italy", "France", "Austria", "Netherlands", "Hungary", "Spain", "Switzerland", "United Kingdom", "France", "Estonia", "Austria", "Sweden","Denmark", "Norway", "Finland")

levels(demog$country.of.residence) = c(levels(demog$country.of.residence), "Other")
demog$country.of.residence[which(!(demog$country.of.residence %in% frequent.countries))] <- "Other"

# frequent nationalities
sort( table(unlist(demog$nationality)),decreasing=TRUE)

demog$nationality = revalue(demog$nationality,
      c("68" = "Germany", "88" = "Italy", "142" = "Poland", "18" = "Belgium", "127" = "Netherlands", "80" = "Hungary", "168" = "Spain", "175" = "Switzerland", "193" = "United Kingdom", "64" = "France", "60" = "Estonia", "11" = "Austria", "132" = "Norway", "143" ="Portugal", "63" = "Finland", "195" = "United States of America", "82" = "India", "186" = "Turkey", "70" = "Greece"))

frequent.nationalities = c("Germany", "Poland", "Belgium", "Italy", "Austria", "Netherlands", "Serbia", "Hungary", "Spain", "Switzerland", "United Kingdom", "France", "Estonia", "Austria", "Norway", "Portugal", "Finland", "United States of America", "India", "Turkey", "Greece")

#68   88  142   18  127   80  168  175   64   60  193   11  195   82  132  147   63  143  186  148   70 
# 174
levels(demog$nationality) = c(levels(demog$nationality), "Other")
demog$nationality[which(!(demog$nationality %in% frequent.nationalities))] <- "Other"

demog$nationality = droplevels(demog$nationality)
demog$country.of.residence = droplevels(demog$country.of.residence)

demog$people.in.household <- factor(demog$people.in.household, levels = c("1", "2", "3", "4", "0"))
demog$quarantine.status <- factor(demog$quarantine.status, levels = c("1", "2", "3", "4", "0"))
demog$household.income.old <- factor(demog$household.income.old, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14"))

# revalue levels for table
demog$gender = revalue(demog$gender, c("1" = "male", "2" = "female", "3" = "diverse"))
demog$current.stay.out.of.town = revalue(demog$current.stay.out.of.town, c("1" = "yes", "2" = "no"))
demog$risk.group = revalue(demog$risk.group, c("1" = "yes", "2" = "no", "3" = "unsure"))
demog$infection.test.status = revalue(demog$infection.test.status, c("0" = "yes", "1" = "no"))
demog$current.stay.out.of.town = revalue(demog$current.stay.out.of.town, c("1" = "yes", "2" = "no"))
demog$quarantine.status = revalue(demog$quarantine.status, c("1" = "none", "2" = "at home", "3" = "in the hospital", "4" = "abroad", "0" = "other"))
demog$relationship.status = revalue(demog$relationship.status, c("1" = "married", "2" = "widowed", "3" = "divorced", "4" = "separated", "5" = "in a domestic partnership or civil union", "6" = "in a steady relationship living together", "7" = "in a steady relationship living apart", "8" = "single", "9" ="other"))
demog$people.in.household = revalue(demog$people.in.household, c("1" = "1", "2" = "2", "3" = "3-4", "4" = "5-6", "0"="more than 6"))
demog$household.income.old = revalue(demog$household.income.old, c("1" = "0-24.999", "2"= "0-4.999", "3" = "5.000-9.999", "4" = "10.000-14.999", "5" = "15.000-24.999", "6" = "25.000-49.999", "7" ="50.000-74.999", "8" = "75.000-99.999", "9" ="100.000-124.999", "11" ="125.000-149.999", "12" ="150.000-174.999", "13" ="175.000-200.000", "14" = "> 200.000"))
demog$survey.language = revalue(demog$survey.language, c("de" = "German", "en" = "English", "es" = "Spanish", "et" = "Estonian", "fr" = "French", "he" = "Hebrew", "hu" = "Hungarian", "it" = "Italian", "nl" = "Dutch", "pl" = "Polish", "sk" = "Slovak", "zh_Hant" = "Chinese", "cs" = "English", "da" = "English", "no" = "English"))

demog$survey.language = droplevels(demog$survey.language)

demog$diagnosed.mental.health = revalue(demog$diagnosed.mental.health, c("0" = "yes", "1" = "no"))
levels(demog$diagnosed.mental.health) = c(levels(demog$diagnosed.mental.health), "not specified")
demog$diagnosed.mental.health[which(demog$diagnosed.mental.health=="")] <- "not specified"
demog$diagnosed.mental.health = droplevels(demog$diagnosed.mental.health)

demog$household.income.old = droplevels(demog$household.income.old)

# occupation
demog$X.0 = if_else(demog$X.0=="1", "Undergoing Education", "")
demog$X = if_else(demog$X=="2", "Education or research", "")
demog$X.1 = if_else(demog$X.1=="3", "Arts, entertainment, sports and media", "")
demog$X.2 = if_else(demog$X.2=="4", "Healthcare", "")
demog$X.3 = if_else(demog$X.3=="5", "First responder (paramedic/firefighter/police)", "")
demog$X.4 = if_else(demog$X.4!="", "Military", "")
demog$X.5 = if_else(demog$X.5!="", "Civil services, politics", "")
demog$X.6 = if_else(demog$X.6!="", "Finance and economy", "")
demog$X.7 = if_else(demog$X.7!="", "Industry", "")
demog$X.8 = if_else(demog$X.8!="", "Sales and services (incl. restaurants and bars)", "")
demog$X.9 = if_else(demog$X.9!="", "Transport (goods and people)", "")
demog$X.10 = if_else(demog$X.10!="", "Installation, maintenance, cleaning, and repairs", "")
demog$X.11 = if_else(demog$X.11!="", "Construction", "")
demog$X.12 = if_else(demog$X.12!="", "Farming, fishing, and forestry", "")
demog$X.13 = if_else(demog$X.13!="", "Office and administrative support", "")
demog$X.14 = if_else(demog$X.14!="", "(Currently) not working", "")
demog$X.15 = if_else(demog$X.15!="", "Other", "")

# occupational status
demog$X.00 = if_else(demog$X.00=="1", "Permanent contract, full-time", "")
demog$X.16 = if_else(demog$X.16=="2", "Temporary contract, full-time", "")
demog$X.17 = if_else(demog$X.17=="3", "Permanent contract, part-time", "")
demog$X.18 = if_else(demog$X.18=="4", "Temporary contract, part-time", "")
demog$X.19 = if_else(demog$X.19=="5", "Self-employed", "")
demog$X.20 = if_else(demog$X.20!="", "Working as a freelancer", "")
demog$X.21 = if_else(demog$X.21!="", "On parental leave", "")
demog$X.22 = if_else(demog$X.22!="", "On sick leave (long term)", "")
demog$X.23 = if_else(demog$X.23!="", "Unemployed on social benefit", "")
demog$X.24 = if_else(demog$X.24!="", "Unemployed without social benefit", "")
demog$X.25 = if_else(demog$X.25!="", "Full-time studying", "")
demog$X.26 = if_else(demog$X.26!="", "Retired", "")
demog$X.27 = if_else(demog$X.27!="", "Other", "")

# group people underage as yes / no
demog$people.in.household.under.18 = if_else(demog$people.in.household.under.18>0, "yes", "no", "missing")
demog$people.in.household.under.18 = as.factor(demog$people.in.household.under.18)

label(demog$age)   <- "Age"
label(demog$gender)   <- "Gender"
label(demog$nationality) <- "Nationality"
label(demog$country.of.residence) <- "Country of residence"
label(demog$risk.group) <- "Belong to a risk group"
label(demog$quarantine.status) <- "Quarantine"
label(demog$infection.test.status) <- "Tested positive for COVID-19"
label(demog$relationship.status) <- "Relationship status"
label(demog$people.in.household) <- "People in household"
label(demog$people.in.household.under.18)   <- "People underage in household"
label(demog$opinion.about.authorities.measures) <- "Agreement with authorities' measures" # Agreement with authorities' measures to curtail spread of the Corona virus
label(demog$adherence.to.recommended.procedures) <- "Following recommended procedures" #Following recommended procedures to limit the spread of the Corona virus
label(demog$current.stay.out.of.town)    <- "Currently out of town"
label(demog$years.of.education) <- "Education"
label(demog$household.income) <- "Household income"
label(demog$diagnosed.mental.health) <- "Diagnosed mental health condition"
label(demog$health.status) <- "Good general health"
label(demog$survey.language) <- "Response language"

label(demog$X.0) <- "Occupation"
label(demog$X) <- ""
label(demog$X.1) <- ""
label(demog$X.2) <- ""
label(demog$X.3) <- ""
label(demog$X.4) <- ""
label(demog$X.5) <- ""
label(demog$X.6) <- ""
label(demog$X.7) <- ""
label(demog$X.8) <- ""
label(demog$X.9) <- ""
label(demog$X.10) <- ""
label(demog$X.11) <- ""
label(demog$X.12) <- ""
label(demog$X.13) <- ""
label(demog$X.14) <- ""
label(demog$X.15) <- ""

label(demog$X.00) <- "Occupational status"
label(demog$X.16) <- ""
label(demog$X.17) <- ""
label(demog$X.18) <- ""
label(demog$X.19) <- ""
label(demog$X.20) <- ""
label(demog$X.21) <- ""
label(demog$X.22) <- ""
label(demog$X.23) <- ""
label(demog$X.24) <- ""
label(demog$X.25) <- ""
label(demog$X.26) <- ""
label(demog$X.27) <- ""

label(demog$years.of.education) <- "Education"
label(demog$household.income.old) <- "Household income"
label(demog$diagnosed.mental.health) <- "Diagnosed mental health condition"
label(demog$health.status) <- "Good general health"

units(demog$age)   <- "years"
units(demog$risk.group)   <- "N"
units(demog$infection.test.status)   <- "N"
units(demog$quarantine.status)   <- "N"
units(demog$relationship.status)   <- "N"
units(demog$years.of.education)    <- "years"
units(demog$people.in.household)   <- "N"
units(demog$people.in.household.under.18)   <- "N"
units(demog$household.income)   <- "€"
units(demog$nationality)   <- "N"
units(demog$country.of.residence)   <- "N"
units(demog$current.stay.out.of.town)   <- "N"
units(demog$health.status)   <- "self-report, 1-5"

my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

t1 = table1(~ age + risk.group + infection.test.status + quarantine.status + opinion.about.authorities.measures + adherence.to.recommended.procedures + diagnosed.mental.health + health.status + relationship.status + people.in.household + people.in.household.under.18 + years.of.education + household.income.old + country.of.residence + current.stay.out.of.town + nationality + survey.language +
              X.0 + X + X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7 + X.8 + X.9 + X.10 + X.11 + X.12 + X.13 + X.14 + X.15 + X.16 + X.00 + X.17 + X.18 + X.19 + X.20 + X.21 + X.22 + X.23 + X.24 + X.25 + X.26 +X.27| gender, data=demog) #, render.continuous=my.render.cont

t1

# describe incomplete cases
data_incomplete = data_en[which(is.na(data_en$complete.eu)),]
length(which(is.na(data_incomplete$Es.SCM)))
length(which(is.na(data_incomplete$Eg.SCM)))
length(which(is.na(data_incomplete$Ec.SCM)))
length(which(is.na(data_incomplete$Es.SSM)))
length(which(is.na(data_incomplete$Eg.SSM)))
length(which(is.na(data_incomplete$Ec.SSM)))

length(which(data_incomplete$missings>0))
length(which(data_incomplete$missings==0&data_incomplete$missing.cov>0))

```

```{r compare model fits}
# stressor count method
xx = !is.na(data_en$complete.eu)# match sample size
# for full sample, use xx = !is.na(data_en$Eg.SCM) & !is.na(data_en$Es.SCM) & !is.na(data_en$Ec.SCM)

m1 <- lm(scale(P)~scale(Eg.SCM),data= data_en[xx,])
m2 <- lm(scale(P)~scale(Es.SCM),data= data_en[xx,])
m3 <- lm(scale(P)~scale(Ec.SCM),data= data_en[xx,])

anova(m1)
anova(m1, m2, m3, test="F") # only for examination - non-nested models cannot be compared

# stressor severity method
xx =  !is.na(data_en$complete.eu)# match sample size
# for full sample, use xx = !is.na(data_en$Eg.SSM) & !is.na(data_en$Es.SSM) & !is.na(data_en$Ec.SSM) 
  
m4 <- lm(scale(P)~scale(Eg.SSM),data= data_en[xx,])
m5 <- lm(scale(P)~scale(Es.SSM),data= data_en[xx,])
m6 <- lm(scale(P)~scale(Ec.SSM),data= data_en[xx,])

# examine explained variance
modelfits = matrix(c(summary(m1)$r.squared, summary(m2)$r.squared, summary(m3)$r.squared, summary(m4)$r.squared, summary(m5)$r.squared, summary(m6)$r.squared), ncol = 1)
rownames(modelfits)  <-  c("Eg.SCM", "Es.SCM", "Ec.SCM", "Eg.SSM", "Es.SSM", "Ec.SSM")
colnames(modelfits) <- c("Rsq")
#stargazer(modelfits, title="R squared for 'P'", type = "text", out="correlation.P.txt")

modelfits.df = data.frame(category = c("Eg.SCM", "Es.SCM", "Ec.SCM", "Eg.SSM", "Es.SSM", "Ec.SSM"),
                          Rsq.adj = c(summary(m1)$adj.r.squared, summary(m2)$adj.r.squared, summary(m3)$adj.r.squared, summary(m4)$adj.r.squared, summary(m5)$adj.r.squared, summary(m6)$adj.r.squared),
                          model = c(as.character(summary(m1)$call)[2], as.character(summary(m2)$call)[2], as.character(summary(m3)$call)[2], as.character(summary(m4)$call)[2], as.character(summary(m5)$call)[2],as.character(summary(m6)$call)[2]))

names(modelfits.df)[2] <- "adj. R-sq."

pandoc.table(modelfits.df, digits = 4, round = 4, keep.trailing.zeros = TRUE, "test")
writeLines(capture.output(pandoc.table(modelfits.df, digits = 4, round = 4, keep.trailing.zeros = TRUE, 'Explained variance by stressor category')), "modelfits.txt")

anova(m6)
anova(m4, m5, m6)
anova(m1, m2, m3, m4, m5, m6)

var(data_en$Es.SCM, na.rm =T)
var(data_en$Eg.SCM, na.rm =T)
```

```{r prepare covariates for Göran, include = FALSE}
names(data_en)[names(data_en) == "X.0"] <- "occupation1"
names(data_en)[names(data_en) == "X"] <- "occupation2"
names(data_en)[names(data_en) == "X.1"] <- "occupation3"
names(data_en)[names(data_en) == "X.2"] <- "occupation4"
names(data_en)[names(data_en) == "X.3"] <- "occupation5"
names(data_en)[names(data_en) == "X.4"] <- "occupation6"
names(data_en)[names(data_en) == "X.5"] <- "occupation7"
names(data_en)[names(data_en) == "X.6"] <- "occupation8"
names(data_en)[names(data_en) == "X.7"] <- "occupation9"
names(data_en)[names(data_en) == "X.8"] <- "occupation10"
names(data_en)[names(data_en) == "X.9"] <- "occupation11"
names(data_en)[names(data_en) == "X.10"] <- "occupation12"
names(data_en)[names(data_en) == "X.11"] <- "occupation13"
names(data_en)[names(data_en) == "X.12"] <- "occupation14"
names(data_en)[names(data_en) == "X.13"] <- "occupation15"
names(data_en)[names(data_en) == "X.14"] <- "occupation16"
names(data_en)[names(data_en) == "X.15"] <- "occupation17"

names(data_en)[names(data_en) == "X.00"] <- "occupational.status1"
names(data_en)[names(data_en) == "X.16"] <- "occupational.status2"
names(data_en)[names(data_en) == "X.17"] <- "occupational.status3"
names(data_en)[names(data_en) == "X.18"] <- "occupational.status4"
names(data_en)[names(data_en) == "X.19"] <- "occupational.status5"
names(data_en)[names(data_en) == "X.20"] <- "occupational.status6"
names(data_en)[names(data_en) == "X.21"] <- "occupational.status7"
names(data_en)[names(data_en) == "X.22"] <- "occupational.status8"
names(data_en)[names(data_en) == "X.23"] <- "occupational.status9"
names(data_en)[names(data_en) == "X.24"] <- "occupational.status10"
names(data_en)[names(data_en) == "X.25"] <- "occupational.status11"
names(data_en)[names(data_en) == "X.26"] <- "occupational.status12"
names(data_en)[names(data_en) == "X.27"] <- "occupational.status13"

which(names(data_en)=="occupation17")
which(names(data_en)=="occupational.status13")

data_en$occupation17[which(nchar(data_en$occupation17)>0)] = "17"
data_en$occupational.status13[which(nchar(data_en$occupational.status13)>0)] = "13"
#test
length(which(data_en$occupation17[which(!is.na(data_en$complete.eu))]=="17"))
length(which(data_en$occupation13[which(!is.na(data_en$complete.eu))]=="13"))
length(which(data_en$occupational.status13[which(!is.na(data_en$complete.eu))]=="13"))
length(which(data_en$occupational.status11[which(!is.na(data_en$complete.eu))]=="11"))

data_en[,c(172:173, 23:38, 40:51)] <- lapply(data_en[,c(172:173, 23:38, 40:51)], as.factor)# convert to factor

data_en$health.status = factor(data_en$health.status, order = TRUE)
data_en$survey.language = revalue(data_en$survey.language, c("de" = "German", "en" = "English", "es" = "Spanish", "et" = "Estonian", "fr" = "French", "he" = "Hebrew", "hu" = "Hungarian", "it" = "Italian", "nl" = "Dutch", "pl" = "Polish", "sk" = "Slovak", "zh_Hant" = "Chinese", "cs" = "English", "da" = "English", "no" = "English"))
data_en$survey.language = droplevels(data_en$survey.language)

covariates = c("age", "gender", "household.income.old","relationship.status", "people.in.household", "people.in.household.under.18","health.status", "risk.group", "infection.test.status", "quarantine.status", "adherence.to.recommended.procedures", "opinion.about.authorities.measures", "years.of.education", "nationality", "Country.of.residence","Survey.language","current.stay.out.of.town", "occupation1", "occupation2", "occupation3", "occupation4", "occupation5", "occupation6", "occupation7", "occupation8", "occupation9", "occupation10", "occupation11", "occupation12", "occupation13", "occupation14", "occupation15", "occupation16", "occupation17", "occupational.status1", "occupational.status2", "occupational.status3", "occupational.status4", "occupational.status5", "occupational.status6", "occupational.status7", "occupational.status8", "occupational.status9", "occupational.status10", "occupational.status11", "occupational.status12", "occupational.status13")

covariates = which(names(data_en) %in% covariates)
data_en_cov = data_en[, which(names(data_en) %in% covariates)]

```

```{r final cleaning}
#### exclude all subjects that were set to NA:
xx = which(is.na(data_en$Respondent.ID))
if(length(xx)>0){data_en = data_en[-xx,]}

# remove unnecessary columns 
xx = grep("X", colnames(data_en))
data_en = data_en[-xx]
# note that in the real data, the above step will also exclude the column IP address

data_en = data_en[, colSums(is.na(data_en)) != nrow(data_en)]

#save(data_en, file = "data_en.RData")
```


```{r old table code - remove?}
# 
# ######### table 1: sample demographics and health status 
# # (all the initial basic variables and people‘s thinking about how the crisis is managed).
# 
# ''' very awkward solution from myside; preparing aver single variable as a vector and rbind in the end to a dataframe.. /mz
# # 
# demovar <- c("age", "gender", "current.stay.out.of.town", "years.of.education", "occupational.status",
#              "household.income", "relationship.status", "people.in.household", "diagnosed.mental.health",
#              "risk.group", "infection.test.status", "quarantine.status", "opinion.about.authorities.measures",
#              "adherence.to.recommended.procedures" )
# demog <- data_en[which(!is.na(data_en$complete.eu)), demovar]
# 
# gender <- 
# demog.summary <- data.frame()
# age_descr <- c("Median age (range)", str_glue("{age[3]}","({age[1]}-{age[6]})"))
# Gender_title <- c("Gender", "")
# Male
# '''
# #demog.summary <- rbind(age_descr,sex_title)
# 
# 
# 
# ######### table 2: average values + SD of the sample in all the dependent and independent variables
# 
# ###step 1: extract all variables of interest into a new dataframe
# ##NOTE: excel sheet SR_Ec.SCM SR_Ec.SSM do not match in R, in R they are defined as SR_c.SCM and SR_c.SSM - if adjusted, change in line below and code above
# variables.of.interest = data_en[, c("SR_Eg.SCM", "SR_Es.SCM", "SR_c.SCM", "SR_Eg.SSM", "SR_Es.SSM", "SR_c.SSM", "PAS", "PSS", "CSS", "OPT", "GSE", "REC", "NEU", "BCS", "PAC", "P", "Es.SCM", "Es.SSM", "Eg.SCM", "Eg.SSM", "Ec.SCM", "Ec.SSM", "CERQSum", "PAS", "age")]
# 
# ###step 2: calculate mean/SD and put into a new dataframe
# ##NOTE: SR... variables are Z-scores, mean of Z-score is always 0 and SD will be 1 -> pointless to calculate?
# Mean = as.data.frame(colMeans(variables.of.interest))
# SD = as.data.frame(apply(variables.of.interest, 2, sd))
# table.mean.sd = cbind(Mean, SD) #combine mean and SD into one table
# colnames(table.mean.sd)[1] <- "Mean" #rename header
# colnames(table.mean.sd)[2] <- "SD" #rename header
# table.mean.sd[, c(1:2)] = table.mean.sd %>% mutate_at(vars(Mean, SD), funs(round(., 2))) #round to two decimals
# 
# ###step 3: use formattable package to create a nice table that can be viewed in the viewer window of R
# formattable(table.mean.sd)
# 
# 
# ######### table 3: intercorrelations of dependent and independent variables
# 
# ###Basic correlation matrix
# corr.matrix = as.data.frame(cor(variables.of.interest, use = "complete.obs")) #intercorrelation matrix
# corr.matrix[, c(1:25)] = corr.matrix %>% mutate_at(vars(c(1:25)), funs(round(., 2))) #round to two decimals
# formattable(corr.matrix, 
#             align = c("c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c") #center values
# )
# 
# ###basic cor(...) does not give p-values, so use rcorr instead (requires more than 4 cases)
# ##variables.of.interest[nrow(variables.of.interest) + 1,] = 1 #testing purpose (needed more than 4 cases, ignore this line) 
# 
# corr.p.matrix <- rcorr(as.matrix(variables.of.interest))
# corr.p.matrix$r #gives correlations
# corr.p.matrix$P #gives p values of correlations
# 
# flattenCorrMatrix <- function(cormat, pmat) { #function to put r and p value into one table
#   ut <- upper.tri(cormat)
#   data.frame(
#     row = rownames(cormat)[row(cormat)[ut]],
#     column = rownames(cormat)[col(cormat)[ut]],
#     cor  =(cormat)[ut],
#     p = pmat[ut]
#   )
# }
# 
# corr.p.matrix <- as.data.frame(flattenCorrMatrix(corr.p.matrix$r, corr.p.matrix$P))
# corr.p.matrix$Sign <- ifelse(corr.p.matrix$p < 0.05, "significant", "not significant") #threshold set at .05
# corr.p.matrix[, c(3,4)] = round(corr.p.matrix[, c(-1, -2, -5)],2) #round 2 decimals, change last value (the 2) if you want more/less decimals
# formattable(corr.p.matrix) #viewer table
# 
# 
# ###helpful intercorrlation matrix graph 
# x = cor(variables.of.interest)
# corrplot(x, type = "lower")
# 
# #

df.cor <- data.frame(PAS=data_en$PAS[which(!is.na(data_en$complete.eu))],
                     PSS=data_en$PSS[which(!is.na(data_en$complete.eu))],
                     CSS=data_en$CSS[which(!is.na(data_en$complete.eu))],
                     OPT=data_en$OPT[which(!is.na(data_en$complete.eu))], 
                     GSE=data_en$GSE[which(!is.na(data_en$complete.eu))], 
                     REC=data_en$REC[which(!is.na(data_en$complete.eu))], 
                     NEU=data_en$NEU[which(!is.na(data_en$complete.eu))], 
                     BCS=data_en$BCS[which(!is.na(data_en$complete.eu))], 
                     PAC=data_en$PAC[which(!is.na(data_en$complete.eu))],
                     R.SSM.C=(data_en$SR_c.SSM[which(!is.na(data_en$complete.eu))])*-1)
matrix.cor <- rcorr(as.matrix(df.cor))

 flattenCorrMatrix <- function(cormat, pmat) { #function to put r and p value into one table
   ut <- upper.tri(cormat)
   data.frame(
     row = rownames(cormat)[row(cormat)[ut]],
     column = rownames(cormat)[col(cormat)[ut]],
     cor  =(cormat)[ut],
     p = pmat[ut]
   )
 }
corr.p.matrix <- as.data.frame(flattenCorrMatrix(matrix.cor$r, matrix.cor$P))

formattable(matrix.cor$r)

df <- as.data.frame(matrix.cor$r)
df <- round(df, 2)
data

#means for diagnosed and non diagnosed cases:

complete.diagnosed <- read.csv("Complete_psych_diagnosis_N1039.csv", header=F)
names(complete.diagnosed)[1] <- "ID"
incomplete.diagnosed <- read.csv("Incomplete_psych_diagnosis_N162.csv", header=F)
names(incomplete.diagnosed)[1] <- "ID"
length(which(data_en$Respondent.ID %in% complete.diagnosed$ID))
#[1] 1039
length(which(!(data_en$Respondent.ID %in% complete.diagnosed$ID)))
#[1] 4826
length(which(data_en$Respondent.ID %in% complete.diagnosed$ID))+length(which(!(data_en$Respondent.ID %in% complete.diagnosed$ID)))
#[1] 5865
length(which(!(data_en$Respondent.ID[data_en$diagnosed.mental.health==""] %in% complete.diagnosed$ID)))
#[1] 589
length(which(data_en$Respondent.ID %in% incomplete.diagnosed$ID))
#[1] 162
length(which(!(data_en$Respondent.ID %in% incomplete.diagnosed$ID)))
#[1] 5703
length(which(data_en$Respondent.ID %in% incomplete.diagnosed$ID))+length(which(!(data_en$Respondent.ID %in% incomplete.diagnosed$ID)))
#[1] 5865
length(which(!(data_en$Respondent.ID[data_en$diagnosed.mental.health==""] %in% incomplete.diagnosed$ID)))
#[1] 589
data_d <- data_en[!data_en$diagnosed.mental.health=="",]
data_d <- data_d[!is.na(data_d$complete.eu),]
mean(data_d$P) #14.9592
mean(data_d$P[which(data_d$Respondent.ID %in% complete.diagnosed$ID)]) # 16.70356
mean(data_d$P[which(!(data_d$Respondent.ID %in% complete.diagnosed$ID))]) #14.43326

sd(data_d$P) #5.729393
sd(data_d$P[which(data_d$Respondent.ID %in% complete.diagnosed$ID)]) # 6.409042
sd(data_d$P[which(!(data_d$Respondent.ID %in% complete.diagnosed$ID))]) #5.399569

data_d_inc <- data_en[!data_en$diagnosed.mental.health=="",]
data_d_inc <- data_d_inc[is.na(data_d_inc$complete.eu),]
mean(data_d_inc$P, na.rm=T) #15.94652
mean(data_d_inc$P[which(data_d_inc$Respondent.ID %in% incomplete.diagnosed$ID)], na.rm=T) #6.73065
mean(data_d_inc$P[which(!(data_d_inc$Respondent.ID %in% incomplete.diagnosed$ID))], na.rm=T) #5.419887

sd(data_d_inc$P, na.rm=T) #5.894808
sd(data_d_inc$P[which(data_d_inc$Respondent.ID %in% incomplete.diagnosed$ID)], na.rm=T) #6.73065
sd(data_d_inc$P[which(!(data_d_inc$Respondent.ID %in% incomplete.diagnosed$ID))], na.rm=T) #5.419887

#table diagnoses:

table.comp <- read.csv("Complete_diagnosis_table.csv", header=F, sep=";", encoding = "UTF-8")
table.incomp <- read.csv("Incomplete_diagnosis_table.csv", header=F, sep=";", encoding="UTF-8")
table.diagnoses <- data.frame(diagnose=table.comp$V1, "complete respondents (N=4485)"=table.comp$V2,"percentage complete"=round(table.comp$V2/4485*100,2), "incomplete respondents (N=791)"=table.incomp$V2, "percentage incomplete"=round(table.incomp$V2/4485*100, 2))
table.diagnoses$diagnose <- as.character(table.diagnoses$diagnose)

table.diagnoses <- rbind(table.diagnoses, c("total", sum(table.diagnoses$complete.respondents..N.4485.), "",sum(table.diagnoses$incomplete.respondents..N.791.),""))

format_table(table.diagnoses)                                        

```

```{r check used languages, include=FALSE}
table(data_en$survey.language)
table(data_en$survey.language[which(!is.na(data_en$complete.eu))])
table(data_en$survey.language[which(is.na(data_en$complete.eu))])

# check dates of survey language
min(data_en$Start.DateTime[which(data_en$survey.language=="cs")])
min(data_en$Start.DateTime[which(data_en$survey.language=="da")])
min(data_en$Start.DateTime[which(data_en$survey.language=="de")])
min(data_en$Start.DateTime[which(data_en$survey.language=="en")])
min(data_en$Start.DateTime[which(data_en$survey.language=="es")])
min(data_en$Start.DateTime[which(data_en$survey.language=="et")])
min(data_en$Start.DateTime[which(data_en$survey.language=="fr")])
min(data_en$Start.DateTime[which(data_en$survey.language=="he")])
min(data_en$Start.DateTime[which(data_en$survey.language=="hu")])
min(data_en$Start.DateTime[which(data_en$survey.language=="it")])
min(data_en$Start.DateTime[which(data_en$survey.language=="nl")])
min(data_en$Start.DateTime[which(data_en$survey.language=="no")])
min(data_en$Start.DateTime[which(data_en$survey.language=="pl")])
min(data_en$Start.DateTime[which(data_en$survey.language=="sk")])
min(data_en$Start.DateTime[which(data_en$survey.language=="zh_Hant")])
```

```{r Matthias is playing around, can be deleted later}

summary(data_en$End.Date)
#                 Min.               1st Qu.                Median                  Mean 
#"2020-03-22 00:00:00" "2020-03-25 00:00:00" "2020-03-29 00:00:00" "2020-03-28 00:28:16" 
#              3rd Qu.                  Max. 
#"2020-03-31 00:00:00" "2020-04-01 00:00:00" 
mean(data_en$CSS[data_en$End.Date<"2020-03-24 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 3.292359
mean(data_en$CSS[data_en$End.Date<="2020-03-25 00:00:00"& data_en$End.Date<"2020-03-28 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 3.240343
mean(data_en$CSS[data_en$End.Date<="2020-03-28 00:00:00"& data_en$End.Date<"2020-03-31 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 3.224191
mean(data_en$CSS[data_en$End.Date<="2020-03-31 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 3.212466
sd(data_en$CSS[data_en$End.Date<"2020-03-24 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 0.6278865
sd(data_en$CSS[data_en$End.Date<="2020-03-25 00:00:00"& data_en$End.Date<"2020-03-28 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 0.6523222
sd(data_en$CSS[data_en$End.Date<="2020-03-31 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 0.6427986
sd(data_en$CSS[data_en$End.Date<="2020-03-28 00:00:00"& data_en$End.Date<"2020-03-31 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 0.639339
length(data_en$CSS[data_en$End.Date<"2020-03-24 00:00:00" & data_en$survey.language=="de"])
#[1] 318
length(data_en$CSS[data_en$End.Date<="2020-03-25 00:00:00"& data_en$End.Date<"2020-03-28 00:00:00" & data_en$survey.language=="de"])
#[1] 983
length(data_en$CSS[data_en$End.Date<="2020-03-28 00:00:00"& data_en$End.Date<"2020-03-31 00:00:00" & data_en$survey.language=="de"])
#[1] 1723
length(data_en$CSS[data_en$End.Date<="2020-03-31 00:00:00" & data_en$survey.language=="de"])
#[1] 2303
mean(data_en$CSS[data_en$End.Date<"2020-03-25 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 3.289076
sd(data_en$CSS[data_en$End.Date<"2020-03-25 00:00:00" & data_en$survey.language=="de"], na.rm=T)
#[1] 0.62802
length(data_en$CSS[data_en$End.Date<"2020-03-25 00:00:00" & data_en$survey.language=="de"])
#[1] 628

